What's in a name? On the automated topic naming of software maintenance activities

### 2010-03-06 icpc/icpc2010.tex

%To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above. We can then categorize development artifacts as belonging to one or more of these topics.

% This correlates issues with topics
As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

%Topics arise from the wide variety of issues which occurs during a software project's lifecycle. These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

@@

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic curve, we will refer to this as \emph{ROC} although it is sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (an \emph{A} is $0.9$, a \emph{C} is $0.6$), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

@@

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to 1 or more documents. Our problem is that these topics are not easy to interpret and we feel that automatic labelling or naming of these topics would be helpful with respect to interpreting what a topic is about. LDA's topics extracted in an unsupervised manner, LDA relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

@@

\subsection{Generating the data}
From each project, we used source control commit comments, these are the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.

From that data-set, we created an XML file which windowed commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied an Blei's LDA implementation against the word distributions of these commits, and generated list of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

@@

Topics are word distributions: essentially lists of words ranked by frequency, thus they can be burdensome to interpret and hard to distinguish and understand.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that a word from our word list. We used several different word lists for comparison.

@@

In conclusion we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics. Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet. We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.

>>>>

% To have meaningful discussions about how software development is progressing, we must devise appropriate labels for our work topics, in contrast to the quote above. We can then categorize development artifacts as belonging to one or more of these topics.

Topics arise from the wide variety of issues which occur during a software project's lifecycle. These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

%As development of a software project progresses, developers must concern themselves with a wide variety of issues --- or ``topics'' --- that relate to the problem domain, the processes and tools used, the development artifacts themselves, etc.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

@@

\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top ten most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (ROC) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

@@

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents. Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear. We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

@@

\subsection{Generating the data}
From each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as common English words \emph{the} and \emph{at}.

From that data-set, we created an XML file which separated commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated list of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

@@

Topics are word distributions: essentially lists of words ranked by frequency, which can be burdensome to interpret and hard to distinguish and understand.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words. We used several different word lists for comparison.

@@

In conclusion, we have demonstrated that static but domain specific knowledge can improve unsupervised labelling of extracted topics. Our \textsf{exp2} experiment used small accurate word bags to label topics but performed just as well as \textsf{exp3}, which used many more general terms from WordNet. We then showed that with some supervision, and by using efficient machine learners based on Naive Bayesian classifiers, we could improve the accuracy of automatic labelling topics even further.
